---
title: "GMM SF1 all tickers"
author: "Robert Thomson"
date: "2026-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GMM on the 1561 observation SF1

Load Packages
```{r}
library(tidyverse)
library(mclust)
library(pheatmap)
```

## Getting the data in, in a way so i dont have to run my data script everytime I change something

```{r}
source("C:\\Users\\Robert\\OneDrive\\Documents\\R Working Directory\\STAT 4051\\Project\\US_equity_data_v4.R")
source("C:\\Users\\Robert\\OneDrive\\Documents\\R Working Directory\\STAT 4051\\Project\\Non_ML_Report_vizs.R")
```

The actual stuff I will cluster
```{r}
SF1_32feat <- SF1_tibble
```
The actual stuff I will cluster with sector and industry
```{r}
SF1_32feat_wLabels <- SF1_tibble_w_labels
```

Checking pairwise feature correlations
```{r}
SF1_32feat_corr <- cor(SF1_32feat)
SF1_32feat_corr <- as_tibble(SF1_32feat_corr)
which( 1 > abs(SF1_32feat_corr) & abs(SF1_32feat_corr) >= .9, arr.ind = T)
```
Going to keep bvps(4) and rel_rnd(25) and drop the rest

```{r}
SF1_27feat <- SF1_32feat %>% select(-roe, -fcfps, -netmargin, -ps, -ebitdamargin)
```

Rechecking correlations
```{r}
SF1_27feat_corr <- cor(SF1_27feat)
which(1 > SF1_27feat_corr & SF1_27feat_corr > .9, arr.ind = T)
which(1 > SF1_27feat_corr & SF1_27feat_corr > .85, arr.ind = T)
which(1 > SF1_27feat_corr & SF1_27feat_corr > .8, arr.ind = T)
```
removing rel_sgna(25)
```{r}
SF1_26feat <- SF1_27feat %>% select(-rel_sgna)
```

Removing marketcap, pb, pe, evebitda, and divyield because they incorporate market prices for the company and I want this to be just financial statement stuff (we have other data for market price stuff)
```{r}
SF1_21feat <- SF1_26feat %>% select(-marketcap, -pb, -pe, -evebitda, -divyield)
```

Checking outliers
```{r}
boxplot(SF1_21feat, las = 2)
```

Still some gigantic outliers, going to try winsorizing at .99 and .01 to see how much it helps. I know it will result in losing some information but I want the clustering to be informative and not completely driven by +40 or -40 SD outliers

Winsorizing the 21 feature SF1 dataset

```{r}
winsorize <- function(x, p = 0.01) {
  q <- quantile(x, c(p, 1-p))

  case_when(x < q[1] ~ q[1], x > q[2] ~ q[2], TRUE ~ x)
}

SF1_21feat_winsor <- sapply(SF1_21feat, winsorize)
boxplot(SF1_21feat_winsor, las=2)
```
Still plenty of outliers but the magnitude of the outliers is much better for a clustering representative of all the features

Making the GMM for the 21 feature SF1 data 
```{r}
gmm_SF1_21_winsor <- Mclust(SF1_21feat_winsor, G=1:20, verbose = T)
```

Peeking at it
```{r}
summary(gmm_SF1_21_winsor)
```
```{r}
summary(gmm_SF1_21_winsor$uncertainty)
```
Heatmap of the clusters(rows) and mean of the observations in each cluster for each feature(columns)
```{r}
pheatmap(t(gmm_SF1_21_winsor$parameters$mean), cluster_rows = F)
```
Plotting the BIC for all models in the Mclust() call
```{r}
plot(gmm_SF1_21_winsor, what = "BIC")
```
Making an object to hold all the cluster assignments for each obsevation in SF1_24feat
```{r}
vec_of_GMMassignments <- as.vector(gmm_SF1_21_winsor$classification)
SF1_21feat_w_GMMclas <- SF1_21feat %>% mutate(ticker = SF1_32feat_wLabels$ticker)
SF1_21feat_w_GMMclas <- SF1_21feat_w_GMMclas %>% mutate(sector = SF1_tibble_w_tickersANDLabels$sector,
      industry = SF1_tibble_w_tickersANDLabels$industry)
SF1_21feat_w_GMMclas$cluster <- vec_of_GMMassignments 
Tickers1561_w_GMMstuff <- SF1_21feat_w_GMMclas %>% select(ticker, sector, industry, cluster)
```

Looking more at the uncertainties for each observation, i.e. which are less certainly in their cluster
```{r}
uncerts <- gmm_SF1_21_winsor$uncertainty
Tickers1561_w_GMMstuff$uncertainty <- uncerts
ggplot(Tickers1561_w_GMMstuff, aes(x=cluster, y=uncertainty, group = cluster))+geom_boxplot()
```

Visualizing the sector and industry representation of each cluster
```{r}
just_sector_and_GMMclust <- Tickers1561_w_GMMstuff %>% select(sector, cluster)
gmm_clust_sector_barchart <- just_sector_and_GMMclust %>% ggplot(aes(x=cluster, fill = sector))+
  geom_bar() +
  scale_fill_manual(values = sector_colors)
gmm_clust_sector_barchart
```

```{r}
#just_sector_and_GMMclust <- ungroup(just_sector_and_GMMclust)
just_sector_and_GMMclust <- just_sector_and_GMMclust %>% group_by(cluster)
just_sector_and_GMMclust %>% filter(cluster == 1) %>% dplyr::count(sector, sort = T)
just_sector_and_GMMclust %>% filter(cluster == 2) %>% dplyr::count(sector, sort = T)
just_sector_and_GMMclust %>% filter(cluster == 3) %>% dplyr::count(sector, sort = T)
just_sector_and_GMMclust %>% filter(cluster == 4) %>% dplyr::count(sector, sort = T)
just_sector_and_GMMclust %>% filter(cluster == 5) %>% dplyr::count(sector, sort = T)
```


















